The Form of Uncertainty Affects Selection for Social Learning

### 2023-02-26 ms.tex
Updates from Overleaf

Uncertainty weighs particularly heavily on human adaptation because of
our long lifespans, cosmopolitan societies, and dispersal across highly variable environments, requiring coarse-grained and plastic behavioral adaptations \citep{levins1962}. However, the term ``uncertainty'' is often used loosely in a way that fails to distinguish it from risk. \emph{Risk} represents a decision with a variable payoff where the state space and probabilities associated with the different outcomes are both known. In contrast, \emph{uncertainty} implies that the outcome probabilities, and possibly the state space, are not known \citep{knight1921, keynes1921}. Furthermore, usage of the term uncertainty often conflates several different interpretations of that word. For example, different models define uncertainty in terms of the rate of environmental change, spatial environmental variation, or the reliability of information acquired from the environment. %Furthermore, in translating mathematical models to verbal predictions, the formal meaning of terms like uncertainty is often lost \citep{knight1921, lawson1988probability, volz2012}, facilitating such conflation. %Furthermore, it can be important to distinguish a narrower definition of uncertainty wherein the probabilities of outcomes remain unknown~\citep{knight1921,volz2012}. 
As the number of formal models of social learning has expanded, an increasing number of modeling choices~\citep{Kendal2018} and formalizations of
uncertainty have made it difficult to compare across models or to consolidate our
understanding of the contexts in which social learning under uncertainty is adaptive. 
% Finally, while risk and uncertainty are often used interchangeably to refer to
% non-deterministic outcomes,

%Other (asocial) forms of learning are also likely adaptations to uncertainty that should be accounted for, but are typically glossed over in models.  
Learning, be it social or not, is an iterative process during which an individual acquires information, forms representations and predictions, and then tests and refines those representations and predictions to manage uncertainty~\citep{jacobs2011bayesian,clark2013whatever}. Because cultural evolutionary models are frequently designed to explain \emph{social} learning mechanisms, they often contrast social learning with a simplistic mechanism for individual learning. In particular, a common assumption is that individual learners can pay a cost to definitively learn the optimal behavior for a particular environment with certainty. This assumption implies that social learning, by observing an individual learner, is a good bet if the environment has not changed. Such modeling choices may lead us to overestimate both the extent to which individual learning provides quality information and the value of social learning in a population of individual learners. % , both for asocially adapting to uncertainty and for social learners to take advantage of.
%and for the individual to adapt to the challenges assumed to be solved by social learning. 
% underestimate the quality of information
% available for social learning.

@@

\paragraph{Selection-set size.}
When one does not know which option to take, uncertainty increases with the number of options, which we call the selection-set size.
% This is a mathematical fact at the
% core of information theory: the number of bits required to reliably encode the optimal solution increases by one every time the
% selection-set doubles in size~\citep{mackay2003information}. 
% This is intuitive; the more options to consider, the greater one's uncertainty about which option is the best choice. 
In many studies of social learning, the selection set is often limited to just two options.  For example, in empirical studies of bumble bees (\emph{Bombus
terresteris}) \citep{Baracchi2018} and great tits (\emph{Parus major}) \citep{Aplin2017}, experimenters provided two behaviors from which the bees or birds
could choose, with one yielding a higher payoff than the other. %respectively,
%each yielding greater or smaller payoffs depending on experimental treatment.
Similarly, many human studies have used just two or three possible
behaviors~\citep{McElreath2005,Morgan2012, Toyokawa2019}. Behavioral study designs are
often motivated by modeling studies with similar formulations~\citep{Rogers1988,boyd1995does,Feldman1996,
perreault2012bayesian}.
Larger, more open sets of behavioral choices are not uncommon in experiments trying to capture more complex or naturalistic tasks~\citep{derex2013, wasielewski2014}, but these map imperfectly onto the theoretical literature that tends to use a limited set of behavioral options.
Some models have studied systems with larger, but defined, selection sets~\citep{Rendell2010, lindstrom2016co}, while in others the number of options is subsumed under the probability that a learner gets the right  answer~\citep{Feldman1996,Enquist2007}.
%\mt{How does this work exactly in Enquist et al 2007?}. 
%Researchers have also compared models with a different number of environment states, corresponding to a different number of optimal behaviors through time~\citep{Feldman1996}. 
Rarely is the size of the selection set explored explicitly as a source of  uncertainty \citep[though see][]{Muthukrishna2016a}, even though the number of options one has is likely to increase the difficulty of the decision task~\citep{haynes2009testing,white2009testing}.  
%These examples share one common feature: the size of the selection set was held static \emph{within} model generations.
% as a measure of changing uncertainty in cultural evolutionary models.

\paragraph{Payoff ambiguity.} Most models of social learning necessarily differentiate between the payoffs for adopting optimal vs. non-optimal behaviors
\citep{BoydRicherson1985,Rogers1988,Enquist2007,Rendell2010,aoki2014evolution}. The size of the difference between these payoffs is usually taken to influence the strength of selection on learning strategies.
However, the ability to discern payoff differences between behaviors is also a source of uncertainty. In reality, payoffs for particular behaviors are not always consistent. A behavior may yield a large payoff sometimes and a small
payoff other times \citep{McElreath2005}. This means that signals about the
relationship between behavior and payoff are often noisy, and differentiating
between behavioral options is in part a problem of signal detection. When the
difference in expected payoffs between optimal and non-optimal behaviors is very
large, this noise matters little, as the signal is still very clear. However, when
the expected payoffs of different behaviors are similar relative to the size of
their variances, ambiguity arises about which behaviors really are superior.
Smaller differences between payoffs corresponds to larger ambiguity. Payoff
ambiguity has been manipulated in both theoretical~\citep{perreault2012bayesian}
and empirical~\citep{McElreath2005,Toyokawa2019} studies, both of which support the
claim that payoff ambiguity increases the reliance on social information.
Importantly, payoff ambiguity %differs from selection-set size in that it not only 
affects both uncertainty \emph{and} the strength of selection, with smaller payoff differences leading to greater uncertainty for a learner and weaker evolutionary selection favoring optimal strategies. 
\edit{It is an outstanding problem to understand how payoff variance and payoff
distributions generally affect evolved behavioral strategies~\citep{Haaland2019}.
For the sake of theoretical clarity for our goal of understanding the
\emph{interaction} of various sources of uncertainty we only operationalize payoff
ambiguity as the difference between expected payoffs, even though empirical studies
have sometimes included both forms of payoff uncertainty~\citep{McElreath2005,Toyokawa2019}, and a more complete model would include differences in mean and variance of payoffs.}

@@

\paragraph{Effective lifespan.} The more opportunities an agent has to learn during its lifespan, the more uncertainty can be reduced, assuming a stationary environment. Correspondingly, a reduction in the number of opportunities to learn will increase the uncertainty about which behavioral options are available and what their associated
payoffs are. %Such uncertainty can be passed on to future generations.
% Learning is an iterative process by which one repeatedly acquires information, forms
% representations and predictions, and tests and refines those representations and
% predictions to manage uncertainty \citep{jacobs2011bayesian,clark2013whatever}. 
We refer to the number of individual learning opportunities within a generation as an individual's effective lifespan to highlight that it is the number of opportunities to learn about the payoffs associated with a behavior, rather than the number of sunrises one experiences, that determines one's uncertainty about the behavioral options.  
%throughout the individual's lifespan that matters here, not necessarily how long they live in the absence of relevant learning opportunities.

Empirically, the number of learning opportunities can be manipulated in the lab, and in the real world will tend to correlate with an individual's relative age. In multi-round studies of information use in novel tasks, US participants' use of social
information declined precipitously across rounds~\citep{McElreath2005}, suggesting
they were more likely to use social information when they were most uncertain about the task early on. In a more naturalistic context, \citet{Aplin2017} found that younger great tits more readily used social information compared to older individuals, possibly because they had accrued less information via individual learning, and possibly because younger individuals have the most to gain (because of higher reproductive value) by switching to superior behavioral options. Cross-cultural studies have highlighted the importance of childhood as a phase of heavy social learning in humans \citep{Reyes2016}. Young children are more likely to
acquire their beliefs and simple skills from their parents than are older children
or adults, which is at least partly due to the differential knowledge accrual
between young children and the older adults to whom they direct the most trust and
attention \citep{kline2013teaching}. While these phenomena map on imperfectly to effective lifespan in our model given social learning is only allowed intergenerationally, effective lifespans for learning do vary across tasks, individuals, and species, yet most models assume only one learning opportunity per generation \citep{BoydRicherson1985, Feldman1996,Henrich1998, perreault2012bayesian}. Some models do allow several cultural generations within one genetic generation \citep{Enquist2007,Rendell2010,lindstrom2016co}, but little formal theory explicitly examines the role of learning opportunities in the evolution of social learning.

@@

While the cultural-evolution literature suggests that several forms of uncertainty have played a role in the evolution of social learning, other cognitive mechanisms have likely also evolved in response to uncertainty \citep{volz2012,johnson2013evolution,van2018uncertainty}. Many of these are flexible learning mechanisms that do not require imitating the behaviors of others. For example, when faced with greater uncertainty individuals may adopt more exploratory learning strategies and may even preferentially test behaviors with greater observed payoff variance \citep{Wilson2014,Gershman2019}.

@@

The structure of the environment is specified by several parameters that remain fixed throughout the course of a given simulation (Table~\ref{tab:uncertaintyParameters}).  The environment affords $B$ behaviors, where $B$ is the \emph{selection-set size}. Uncertainty increases with $B$. Each behavior is indexed by $b$ and yields a payoff of 1 with probability $\pi_b$ and a payoff of 0 otherwise. As such, $\pi_b$ is equivalently the expected payoff of behavior $b$. At the beginning of each generation, one behavior is optimal and yields expected payoff $\pihigh$; the other $B-1$ behaviors yield a lower payoff, $\pilow$. The \emph{payoff ambiguity} is defined by $\pihigh - \pilow$. We fix $\pihigh = 0.9$ to be constant across all simulations, so payoff ambiguity is varied by changing the value of $\pilow$. In other words, payoff ambiguity, and therefore uncertainty, increases with $\pilow$.  Agents have $L$ time steps to learn and act per generation (i.e., the \emph{effective lifespan}). Uncertainty decreases with $L$. At the start of each generation a new optimal behavior is selected at random from the set of all possible behaviors with probability $u$ (the \emph{environmental variability}), otherwise the optimal behavior remains unchanged from the previous generation. Uncertainty increases with $u$.

\vspace{2em}
\begin{table}[h]
\caption{Environmental Parameters. These include our four main uncertainty parameters under investigation; 
$\pilow$, $B$, $L$, and $u$. Bold indicates default value tested.} %\emph{default value tested}.}
    \label{tab:uncertaintyParameters}
    \centering %\hspace{-3em}
    \begin{tabular}{cp{4.0in}p{1.25in}} \toprule

@@

\begin{figure}
  \caption{Agents are randomly initialized as social learners or not, with their
  payoff observations all initialized to zero (A). Then agents begin selecting
and performing behaviors and accumulating payoffs, which goes on for $L$
timesteps (B). After $L$ time steps, agents are selected to reproduce,
social learner children learn from a member of their parent's generation, and
the previous generation dies off (C). The simulation stops if children are all
social or asocial learners (i.e.\ the system reaches fixation).
Otherwise it repeats another generation and evolution continues.} 
% \cm{What does (3) denote at end of Fig 1 caption? Does having only green agents in Fig1B give the impression that only social learners are learning within lifetime? Could the agent be half green/purple? Is it weird that in Fig1C a teacher with relatively low payoffs is being selected? Could be fixed by having the dotted line go from rich purple women's head to green baby's. }
  \label{fig:schematic}
  \centering
    \includegraphics[width=1.05\textwidth]{Figures/IntraInterGenerationalDynamics.pdf}
\end{figure}

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values %(Table~\ref{tab:modelParameters}) 
and observed how frequently social learning evolved across 1000 trial simulations for each combination of parameters tested. %We varied $u \in \{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  
\edit{We observed three outcome measures: first, the social learning (SL) fixation ferquency, denoted
$\meansl$, is how frequently
social learning evolved to fixation across all trials for a given parameter setting. To support our
analysis about why we observe the patterns in social learning fixation frequency that we did, we also
measured the mean number of generations to fixation, which measures the strength of selection.} Finally we
measured the average net payoffs at the end of the simulation across agents and trials, normalized by lifespan. We analyze
outcomes by plotting $\meansl$ on the y-axis and environmental variability on
x-axis, since we theoretically expect that $\meansl$ will always decrease monotonically from 1 to 0 as $u$ increases. %Note that the hypothesis-testing concept of significance is meaningless here because we could make any small outcome difference ``significant'' by running more simulation trials.  We instead study the patterns of outcome variables over systematically varied uncertainty factor values.

\begin{table}[h]
    \caption{Outcome Variables. All averages are computed across trials at the end of the last generation.}
    \label{tab:outcomeVariables}
    \centering %\hspace{-3em}
    \begin{tabular}{clp{0.85in}} \toprule

@@

The number of behavioral options, i.e., larger $B$, favors the evolution of social learning. Increasing the selection-set size
expands the range of $u$ over which social learning is favored for two reasons. First, when the environment remains constant between
generations, the value of social learning increases with more behavioral options, because it effectively increases the number of
observations that one can learn from by comparing the payoffs of several models. In other words, relying \emph{only} on one's
individual trial-and-error learning is less likely to yield useful information when the search space is
large. Second, when the environment changes, bias against the new optimal is weaker with more behavioral
options. For example, with only two behavioral choices ($B = 2$), a common assumption in many models, social
learning is particularly detrimental when the environment changes since more of the population will have
arrived at the optimal behavior in the previous generation and socially learning from them will be biased
against the correct choice. These two mechanisms make it particularly profitable to engage in social
learning when the selection-set size is larger---for large values of $B$ social learning can even be favored
when the environment is more likely to change than not (i.e.,
$u>0.5$\edit{---see Figure~\ref{fig:slCeiling}}).

@@

\begin{figure}
  \caption{Social learning fixation frequency across trials (y-axes) monotonically decreases as 
  environmental variability, $u$, increases (x-axes) in most uncertainty contexts. 
  Other uncertainty values payoff ambiguity, $\pilow$, (rows), number of behavioral options, $B$, (columns), and effective lifespan, $L$, (keys) shift and flatten the slope from all-social-learner populations to all-asocial-learner
  populations.}
  \label{fig:mainResults}
  \centering
    \includegraphics[width=\textwidth]{Figures/mainResultsPlots.pdf}
\end{figure}

\begin{figure}
  \caption{As the selection-set size, $B$, and effective lifespan, $L$, increase (set equal along x-axis), 
  social learning frequently evolves across an increasingly large range of
  environmental variability values, i.e., up to
  increasingly large values of the social learning ceiling
(y-axis; defined in Equation~\ref{eq:slCeiling}).}
  \label{fig:slCeiling}
  \centering
    \includegraphics[width=0.8\textwidth]{Figures/SL_ceiling.pdf}
\end{figure}

@@

When payoff ambiguity and selection-set size was small ($\pilow=0.1$ and $B=2$), and lifespan was long ($L \geq 4$), asocial learning fixated as a hedge against periods of protracted environmental instability, even though expected
homogeneous social payoffs were greater than homogeneous asocial payoffs (Figure~\ref{fig:payoffTimeseries}A). Runs of environmental change strongly affected the geometric mean fitness of social learners who were weighing
outdated information too heavily. \edit{In other words, environmental change lowers social learners' mean payoffs, but social learner payoffs build back up as the information quality in the population improves (blue dots, Figure~\ref{fig:payoffTimeseries}A).  Even though an homogenous population of social learners would outperform individual learners in this case, unpredictable environmental stochasticity undermines the reliability of social learning.  When lifespan was minimal ($L=1$) and environmental variability was moderate ($u\approx0.5$), the opposite, unexpected outcome could occur where social learning fixates despite the fact that an homogenous population of asocial learners would outperform an homogenous population of social learners (Figure~\ref{fig:payoffTimeseries}B). In this case, it takes relatively few consecutive generations with environmental stability to attract the population to all become social learners.}

@@

Our model necessarily made simplifying assumptions. These were chosen judiciously to
address the main question of how uncertainty affects the evolution of social
learning. Two critical assumptions we made have important consequences for comparing
our results with previous findings in the literature. First, we assumed all agents
engage in individual learning. This means that individual learning is effectively
free, though potentially inaccurate. Although producing new knowledge may be costly,
many other adaptive problems take the form of essentially cost-free individual
learning. For example, it is hard not to encode association between clouds and rain,
between thirst and drinking water, or between friendship and warm laughter.
\edit{If there were an exogenous cost of individual learning, this would likely
  increase the social learning ceiling across uncertainty parameter settings.
Our model operationalizes environmental variability as fully stochastic,
which forces evolution towards either a population of social or asocial learners
since environmental change is unpredictable, unlike precursor
models that allowed for persistent frequency
dependence~\citep{BoydRicherson1985,Rogers1988,Feldman1996}.} Second, we only allow for
intergenerational social learning and environmental change. Within generations,
agents can only learn individually. This assumption \edit{is consistent with previous modeling work starting at least from \cite{Rogers1988}, and therefore} allows us to replicate the
classic finding that if the environment changes too quickly, socially-learned
information may become \edit{outdated. Adding horizontal social learning between
vertical/oblique learning in this model would likely expand the range of
environmental variability over which social learning would
evolve~\citep{Turner2022}, i.e., horizontal learning would increase what we have
called the social learning ceiling. However, depending on the
setting, there may be little difference between horizontal and oblique learning, for
example if the generations are purely cultural ‘generations’ where parents, and
others of the parent's generation, may also be peers.}

>>>>

Uncertainty weighs particularly heavily on human adaptation because of
our long lifespans, cosmopolitan distribution, and dispersal across highly variable environments, requiring coarse-grained and plastic behavioral adaptations \citep{levins1962}. However, the term ``uncertainty'' is often used loosely in a way that fails to distinguish it from risk. \emph{Risk} represents a decision with a variable payoff where the state space and probabilities associated with the different outcomes are both known. In contrast, \emph{uncertainty} implies that the outcome probabilities, and possibly the state space, are not known \citep{knight1921, keynes1921}. Furthermore, usage of the term uncertainty often conflates several different interpretations of that word. For example, different models define uncertainty in terms of the rate of environmental change, spatial environmental variation, or the reliability of information acquired from the environment. %Furthermore, in translating mathematical models to verbal predictions, the formal meaning of terms like uncertainty is often lost \citep{knight1921, lawson1988probability, volz2012}, facilitating such conflation. %Furthermore, it can be important to distinguish a narrower definition of uncertainty wherein the probabilities of outcomes remain unknown~\citep{knight1921,volz2012}. 
As the number of formal models of social learning has expanded, an increasing number of modeling choices~\citep{Kendal2018} and formalizations of
uncertainty have made it difficult to compare across models or to consolidate our
understanding of the contexts in which social learning under uncertainty is adaptive. 
% Finally, while risk and uncertainty are often used interchangeably to refer to
% non-deterministic outcomes,

%Other (asocial) forms of learning are also likely adaptations to uncertainty that should be accounted for, but are typically glossed over in models.  
Learning, social or otherwise, is an iterative process during which an individual acquires information, forms representations and predictions, and then tests and refines those representations and predictions to manage uncertainty~\citep{jacobs2011bayesian,clark2013whatever}. Because cultural evolutionary models are frequently designed to explain \emph{social} learning mechanisms, they often contrast social learning with a simplistic mechanism for individual learning. In particular, a common assumption is that individual learners can pay a cost to definitively learn the optimal behavior for a particular environment with certainty. This assumption implies that social learning, by observing an individual learner, is a good bet if the environment has not changed. Such modeling choices may lead us to overestimate both the extent to which individual learning provides quality information and the value of social learning in a population of individual learners. % , both for asocially adapting to uncertainty and for social learners to take advantage of.
%and for the individual to adapt to the challenges assumed to be solved by social learning. 
% underestimate the quality of information
% available for social learning.

@@

\paragraph{Selection-set size.}
When one does not know which option to take, uncertainty increases with the number of options, which we call the selection-set size.
% This is a mathematical fact at the
% core of information theory: the number of bits required to reliably encode the optimal solution increases by one every time the
% selection-set doubles in size~\citep{mackay2003information}. 
% This is intuitive; the more options to consider, the greater one's uncertainty about which option is the best choice. 
In many studies of social learning, the selection set is often limited to just two options.  For example, in empirical studies of bumble bees (\emph{Bombus
terresteris}) \citep{Baracchi2018} and great tits (\emph{Parus major}) \citep{Aplin2017}, experimenters provided two behaviors from which the bees or birds
could choose, with one yielding a higher payoff than the other. %respectively,
%each yielding greater or smaller payoffs depending on experimental treatment.
Similarly, many human studies have used just two or three possible
behaviors~\citep{McElreath2005,Morgan2012, Toyokawa2019}. The designs of behavioral experiments are 
often motivated by modeling studies with similar formulations~\citep{Rogers1988,boyd1995does,Feldman1996,
perreault2012bayesian}.
Larger, more open sets of behavioral choices are not uncommon in experiments trying to capture more complex or naturalistic tasks~\citep{derex2013, wasielewski2014}, but these map imperfectly onto the theoretical literature that tends to use a limited set of behavioral options.
Some models have studied systems with larger, but defined, selection sets~\citep{Rendell2010, lindstrom2016co}, while in others the number of options is subsumed under the probability that a learner gets the right  answer~\citep{Feldman1996,Enquist2007}.
%\mt{How does this work exactly in Enquist et al 2007?}. 
%Researchers have also compared models with a different number of environment states, corresponding to a different number of optimal behaviors through time~\citep{Feldman1996}. 
Rarely is the size of the selection set explored explicitly as a source of  uncertainty \citep[though see][]{Muthukrishna2016a}, even though the number of options one has is likely to increase the difficulty of the decision task~\citep{haynes2009testing,white2009testing}.  
%These examples share one common feature: the size of the selection set was held static \emph{within} model generations.
% as a measure of changing uncertainty in cultural evolutionary models.

\paragraph{Payoff ambiguity.} Most models of social learning necessarily differentiate between the payoffs for adopting optimal vs. non-optimal behaviors
\citep{BoydRicherson1985,Rogers1988,Enquist2007,Rendell2010,aoki2014evolution}. The size of the difference between these payoffs is usually taken to influence the strength of selection on learning strategies.
However, the ability to discern payoff differences between behaviors is also a source of uncertainty. In reality, payoffs for particular behaviors are not always consistent. A behavior may yield a large payoff sometimes and a small
payoff other times \citep{McElreath2005}. This means that signals about the
relationship between behavior and payoff are often noisy, and differentiating
between behavioral options is in part a problem of signal detection. When the
difference in expected payoffs between optimal and non-optimal behaviors is very
large, this noise matters little, as the signal is still very clear. However, when
the expected payoffs of different behaviors are similar relative to the size of
their variances, ambiguity arises about which behaviors really are superior.
Smaller differences between payoffs corresponds to larger ambiguity. Payoff
ambiguity has been manipulated in both theoretical~\citep{perreault2012bayesian}
and empirical~\citep{McElreath2005,Toyokawa2019} studies, both of which support the
claim that payoff ambiguity increases the reliance on social information.
Importantly, payoff ambiguity %differs from selection-set size in that it not only 
affects both uncertainty \emph{and} the strength of selection, with smaller payoff differences leading to greater uncertainty for a learner and weaker evolutionary selection favoring optimal strategies. 
\edit{It is an outstanding problem to understand how payoff variance and payoff
distributions generally affect evolved behavioral strategies~\citep{Haaland2019}.
For the sake of theoretical clarity for our goal of understanding the
\emph{interaction} of various sources of uncertainty, we only operationalize payoff
ambiguity as the difference between expected payoffs, even though empirical studies
have sometimes included both forms of payoff uncertainty~\citep{McElreath2005,Toyokawa2019}, and a more complete model would include differences in mean and variance of payoffs.}

@@

\paragraph{Effective lifespan.} The more opportunities an agent has to learn during its lifespan, the more uncertainty can be reduced, assuming a stationary environment. Correspondingly, a reduction in the number of opportunities to learn will increase the uncertainty about which behavioral options are available and what their associated
payoffs are. %Such uncertainty can be passed on to future generations.
% Learning is an iterative process by which one repeatedly acquires information, forms
% representations and predictions, and tests and refines those representations and
% predictions to manage uncertainty \citep{jacobs2011bayesian,clark2013whatever}. 
We refer to the number of individual learning opportunities within a generation as an individual's \emph{effective} lifespan to highlight that it is the number of opportunities to learn about the payoffs associated with a behavior, rather than the number of sunrises one experiences, that determines one's uncertainty about the behavioral options.  
%throughout the individual's lifespan that matters here, not necessarily how long they live in the absence of relevant learning opportunities.

Empirically, the number of learning opportunities can be manipulated in the lab and in the real world will tend to correlate with an individual's relative age. In multi-round studies of information use in novel tasks, US participants' use of social
information declined precipitously across rounds~\citep{McElreath2005}, suggesting
they were more likely to use social information when they were most uncertain about the task early on. In a more naturalistic context, \citet{Aplin2017} found that younger great tits more readily used social information compared to older individuals, possibly because they had accrued less information via individual learning, and possibly because younger individuals have the most to gain (because of higher reproductive value) by switching to superior behavioral options. Cross-cultural studies have highlighted the importance of childhood as a phase of heavy social learning in humans \citep{Reyes2016}. Young children are more likely to
acquire their beliefs and simple skills from their parents than are older children
or adults, which is at least partly due to the differential knowledge accrual
between young children and the older adults to whom they direct the most trust and
attention \citep{kline2013teaching}. While these phenomena map imperfectly onto effective lifespan in our model, given social learning is only allowed intergenerationally, \edit{the relative number of individual to social learning opportunities, i.e., this model's version of }effective lifespan, do vary across tasks, individuals, and species, yet most models assume only one learning opportunity per generation \citep{BoydRicherson1985, Feldman1996,Henrich1998, perreault2012bayesian}. Some models do allow several cultural generations within one genetic generation \citep{Enquist2007,Rendell2010,lindstrom2016co}, but little formal theory explicitly examines the role of learning opportunities in the evolution of social learning.

@@

While the cultural-evolution literature suggests that several forms of uncertainty have played a role in the evolution of social learning, other cognitive mechanisms have likely also evolved in response to uncertainty \citep{volz2012,johnson2013evolution,van2018uncertainty}. Many of these are flexible learning mechanisms that do not require imitating the behaviors of others. For example, when faced with greater uncertainty, individuals may adopt more exploratory learning strategies and may even preferentially test behaviors with greater observed payoff variance \citep{Wilson2014,Gershman2019}.

@@

The structure of the environment is specified by several parameters that remain fixed throughout the course of a given simulation (Table~\ref{tab:uncertaintyParameters}).  The environment affords $B$ behaviors, where $B$ is the \emph{selection-set size}. Uncertainty increases with $B$. Each behavior is indexed by $b$ and yields a payoff of 1 with probability $\pi_b$ and a payoff of 0 otherwise. As such, $\pi_b$ is equivalently the expected payoff of behavior $b$. At the beginning of each generation, one behavior is optimal and yields expected payoff $\pihigh$; the other $B-1$ behaviors yield a lower payoff, $\pilow$. The \emph{payoff ambiguity} is defined by $\pihigh - \pilow$. We fix $\pihigh = 0.9$ to be constant across all simulations, so payoff ambiguity is varied by changing the value of $\pilow$. In other words, payoff ambiguity, and therefore uncertainty, increases with $\pilow$.  Agents have $L$ time steps to \edit{act and learn asocially} per generation (i.e., the \emph{effective lifespan}). Uncertainty decreases with $L$. At the start of each generation a new optimal behavior is selected at random from the set of all possible behaviors with probability $u$ (the \emph{environmental variability}), otherwise the optimal behavior remains unchanged from the previous generation. Uncertainty increases with $u$.

\vspace{2em}
\begin{table}[h]
\caption{Environmental parameters. These include our four main uncertainty parameters under investigation; 
$\pilow$, $B$, $L$, and $u$. Bold indicates default value tested.} %\emph{default value tested}.}
    \label{tab:uncertaintyParameters}
    \centering %\hspace{-3em}
    \begin{tabular}{cp{4.0in}p{1.25in}} \toprule

@@

\begin{figure}
  \caption{\edit{Model summary}. Agents are randomly initialized as social learners or not, with their
  payoff observations all initialized to zero (A). Then agents begin selecting
and performing behaviors and accumulating payoffs, which goes on for $L$
timesteps (B). After $L$ time steps, agents are selected to reproduce,
social learner children learn from a member of their parent's generation, and
the previous generation dies off (C). The simulation stops if children are all
social or asocial learners (i.e.\ the system reaches fixation).
Otherwise it repeats another generation and evolution continues.} 
% \cm{What does (3) denote at end of Fig 1 caption? Does having only green agents in Fig1B give the impression that only social learners are learning within lifetime? Could the agent be half green/purple? Is it weird that in Fig1C a teacher with relatively low payoffs is being selected? Could be fixed by having the dotted line go from rich purple women's head to green baby's. }
  \label{fig:schematic}
  \centering
    \includegraphics[width=1.05\textwidth]{Figures/IntraInterGenerationalDynamics.pdf}
\end{figure}

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values %(Table~\ref{tab:modelParameters}) 
and observed how frequently social learning evolved across 1000 trial simulations for each combination of parameters tested. %We varied $u \in \{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  
\edit{We observed three outcome measures: first, the social learning (SL) fixation frequency, denoted
$\meansl$, is how frequently
social learning evolved to fixation across all trials for a given parameter setting. To support our
analysis about why we observe the patterns in social learning fixation frequency that we did, we also
measured the mean number of generations to fixation, which measures the strength of selection.} Finally we
measured the average net payoffs at the end of the simulation across agents and trials, normalized by lifespan. We analyze
outcomes by plotting $\meansl$ on the y-axis and environmental variability on
x-axis, since we theoretically expect that $\meansl$ will always decrease monotonically from 1 to 0 as $u$ increases. %Note that the hypothesis-testing concept of significance is meaningless here because we could make any small outcome difference ``significant'' by running more simulation trials.  We instead study the patterns of outcome variables over systematically varied uncertainty factor values.

\begin{table}[h]
    \caption{Outcome variables. All averages are computed across trials at the end of the last generation.}
    \label{tab:outcomeVariables}
    \centering %\hspace{-3em}
    \begin{tabular}{clp{0.85in}} \toprule

@@

The number of behavioral options, i.e., larger $B$, favors the evolution of social learning. Increasing the selection-set size
expands the range of $u$ over which social learning is favored for two reasons. First, when the environment remains constant between
generations, the value of social learning increases with more behavioral options, because it effectively increases the number of
observations that one can learn from by comparing the payoffs of several models. In other words, relying \emph{only} on one's
individual trial-and-error learning is less likely to yield useful information when the search space is
large. Second, when the environment changes, bias against the new optimal is weaker with more behavioral
options. For example, with only two behavioral choices ($B = 2$), a common assumption in many models, social
learning is particularly detrimental when the environment changes, since more of the population will have
arrived at the optimal behavior in the previous generation and socially learning from them will be biased
against the correct choice. These two mechanisms make it particularly profitable to engage in social
learning when the selection-set size is larger---for large values of $B$ social learning can even be favored
when the environment is more likely to change than not (i.e.,
$u>0.5$\edit{---see Figure~\ref{fig:slCeiling}}).

@@

\begin{figure}
  \caption{\edit{Effect of four forms of uncertainty on evolution of social learning}. Social learning fixation frequency across trials (y-axes) monotonically decreases as 
  environmental variability, $u$, increases (x-axes) in most uncertainty contexts. 
  Other uncertainty values payoff ambiguity, $\pilow$, (rows), number of behavioral options, $B$, (columns), and effective lifespan, $L$, (keys) shift and flatten the slope from all-social-learner populations to all-asocial-learner
  populations.}
  \label{fig:mainResults}
  \centering
    \includegraphics[width=\textwidth]{Figures/mainResultsPlots.pdf}
\end{figure}

\begin{figure}
  \caption{\edit{Effect of uncertainty on social learning ceiling.} As the selection-set size, $B$, and effective lifespan, $L$, increase (set equal along x-axis), 
  social learning frequently evolves across an increasingly large range of
  environmental variability values, i.e., up to
  increasingly large values of the social learning ceiling
(y-axis; defined in Equation~\ref{eq:slCeiling}).}
  \label{fig:slCeiling}
  \centering
    \includegraphics[width=0.8\textwidth]{Figures/SL_ceiling.pdf}
\end{figure}

@@

When payoff ambiguity and selection-set size were small ($\pilow=0.1$ and $B=2$), and lifespan was long ($L \geq 4$), asocial learning fixated as a hedge against periods of protracted environmental instability, even though expected
homogeneous social payoffs were greater than homogeneous asocial payoffs (Figure~\ref{fig:payoffTimeseries}A). Runs of environmental change strongly affected the geometric mean fitness of social learners who were weighing
outdated information too heavily. \edit{In other words, environmental change lowers social learners' mean payoffs, but social-learner payoffs build back up as the information quality in the population improves (blue dots, Figure~\ref{fig:payoffTimeseries}A).  Even though a homogenous population of social learners would outperform individual learners in this case, unpredictable environmental stochasticity undermines the reliability of social learning.  When lifespan was minimal ($L=1$) and environmental variability was moderate ($u\approx0.5$), the opposite, unexpected outcome could occur where social learning fixates despite the fact that a homogenous population of asocial learners would outperform a homogenous population of social learners (Figure~\ref{fig:payoffTimeseries}B). In this case, it takes relatively few consecutive generations with environmental stability to attract the population to all become social learners.}

@@

Our model necessarily made simplifying assumptions. These were chosen judiciously to
address the main question of how uncertainty affects the evolution of social
learning. Two critical assumptions we made have important consequences for comparing
our results with previous findings in the literature. First, we assumed all agents
engage in individual learning. This means that individual learning is effectively
free, though potentially inaccurate. Although producing new knowledge may be costly,
many other adaptive problems take the form of essentially cost-free individual
learning. For example, it is hard not to encode association between clouds and rain,
between thirst and drinking water, or between friendship and warm laughter.
\edit{If there were an exogenous cost of individual learning, this would likely
  increase the social learning ceiling across uncertainty parameter settings.
This means that our model does not capture the information-scrounging rewards to individual learning that produce mixed equilibria in precursor models \cite{Rogers1988}. Furthermore, our model operationalizes environmental variability as fully stochastic in a finite population. This makes it exceedingly unlikely for a heterogeneous population to persist since chance runs of very stable environments will lead to the extinction of asocial learners, while runs of very unstable environments will lead to the extinction of social learners. This further differentiates our results from precursor
models that allowed for persistent frequency
dependence~\citep{BoydRicherson1985,Rogers1988,Feldman1996}.} Second, we only allow for
intergenerational social learning and environmental change. Within generations,
agents can only learn individually. This assumption \edit{is consistent with previous modeling work starting at least from \cite{Rogers1988}, and therefore} allows us to replicate the
classic finding that if the environment changes too quickly, socially-learned
information may become \edit{outdated. Adding horizontal social learning between
vertical/oblique learning in this model would likely expand the range of
environmental variability over which social learning would
evolve~\citep{Turner2022}, i.e., horizontal learning would increase what we have
called the social learning ceiling. However, depending on the
setting, there may be little difference between horizontal and oblique learning, for
example if the generations are purely cultural ‘generations’ where parents, and
others of the parent's generation, may also be peers.}
